---
title: "R Notebook"
output: html_notebook
---

# Based On Course 'Introduction to Time Series Analysis' on Data Camp
# Chapter 1 , Exploratory time series data analysis
#   https://www.datacamp.com/courses/introduction-to-time-series-analysisLinks 



```{r}
unempData =read.csv(url("https://data.ny.gov/api/views/5hyu-bdh8/rows.csv?accessType=DOWNLOAD"),header=TRUE, stringsAsFactors = TRUE)

# Fixing the Date into a YYYY-MM format
unempData$Date <- as.Date(paste(unempData$Year, unempData$Month, sep="-"), "%Y-%M")
# Droping the redundant Year and month 
unempData <- subset(unempData, select=-c(Year,Month))
# odering the Data
unempData <- unempData[with(unempData,order(Area,Date)),]

summary(unempData)
```
```{r}
tail(unempData)
```


Creating Time Series Data for Albany City
```{r}
AC_DATA <- subset(unempData, Area=='Albany City')
AC_DATA_UER <- select(AC_DATA,Unemployment.Rate)
AC_DATA_U <- select(AC_DATA,Unemployed)
AC_DATA_E <- select(AC_DATA,Employed)
AC_DATA_LF <- select(AC_DATA,Labor.Force)

ts_AC<-ts(AC_DATA,start=c(1990,1),end=c(2019,2),frequency=12)
ts_AC_UER<-ts(AC_DATA_UER,start=c(1990,1),end=c(2019,2),frequency=12)
ts_AC_U<-ts(AC_DATA_U,start=c(1990,1),end=c(2019,2),frequency=12)
ts_AC_E<-ts(AC_DATA_E,start=c(1990,1),end=c(2019,2),frequency=12)
ts_AC_LF<-ts(AC_DATA_LF,start=c(1990,1),end=c(2019,2),frequency=12)


library(ggplot2)
head(ts_AC_UER)

```

Les Up Understnad the Albany City Row Data ie AC_DATA and corresponding time series data ts_AC

```{r}
print(" Length ")
length(ts_AC)
```

```{r}


ggplot(data = AC_DATA, aes(x = Date, y = Unemployment.Rate)) +geom_line( color = "#FC4E07")


ggplot(data = AC_DATA, aes(x = Date, y = Unemployment.Rate)) +geom_line(aes(size = Unemployment.Rate/Labor.Force), color = "#FC4E07")
```


We could see some trems in Row Data
Lets See what best a Time Series Object can offer


```{r}
plot(ts_AC)
plot(ts_AC_UER)

```



This data is Equlay spaced with out any missing value
That could also be established from code 
```{r}
print("Start")
start(ts_AC)
print("end")
end(ts_AC)

print("deltat")
# function returns the fixed time interval between observations
deltat(ts_AC)

print("frequency")
#function returns the number of observations per unit time
frequency(ts_AC)

print("time")
time(ts_AC)

print("cycle")
#cycle() function returns the position in the cycle of each observation
cycle(ts_AC)
```

```{r}
print("The Total Missing Values are")
sum( is.na( ts_AC ) )

```

As We See there are No Missing value.. We need not impute it

As we have seen some trend, in the time sesies, some seasonal factor also. We can now try out diff to remove the trend

```{r}
diff_TSAC_UER <- diff(ts_AC_UER,lag = 4)
  
# Plot dz
ts.plot(diff_TSAC_UER)  

# View the length of z and dz, respectively
length(diff_TSAC)
length(ts_AC)

```

We have lost some 24 rows on applying Diff since we have a Lag of 4.
Now all what we see is some white noise.


Lets now fit the Time series data to Arima model and find the Mean and variance

```{r}
diff_TSAC_UER
```

```{r}
# Fit the WN model to y using the arima command
UER_model <- arima(diff_TSAC_UER, order = c(0, 0, 0))
mean(diff_TSAC_UER)
var(diff_TSAC_UER)
print("Model")
UER_model
```


Lets Plot and see how does our Model fits.

```{r}
ts.plot(ts_AC_UER)
modelCoef <- UER_model$coef
abline(0, modelCoef)
```

Lets Do that with another variable, Employed

```{r}
ggplot(data = AC_DATA, aes(x = Date, y = Employed)) +geom_line( color = "#FC4E07")

ggplot(data = AC_DATA, aes(x = Date, y = Employed)) +geom_line(aes(size = Employed/Labor.Force), color = "#FC4E07")

```


```{r}
plot(ts_AC)
plot(ts_AC_LF)
```

Some thing went wrong in year 2000 in our data set. Lets Look further.


```{r}
diff_TSAC_LF <- diff(ts_AC_LF,lag = 4)
ts.plot(diff_TSAC_LF)  
length(diff_TSAC_LF)
length(ts_AC)
```

```{r}
LF_model <- arima(diff_TSAC_LF, order = c(0, 0, 0))
mean(diff_TSAC_LF)
var(diff_TSAC_LF)
print("Model_LF")
LF_model
```

```{r}
ts.plot(ts_AC_LF)
modelCoef_LF <- LF_model$coef
abline(0, modelCoef_LF)
```



When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
